Dynamics of international giving: Exploring individual donor preferences

### 2020-06-23 manuscript/manuscript.Rmd
Suparna edits

---
title: "Dynamics of international giving: Exploring individual donor preferences"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Lewis \& Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  Government restrictions on NGOs have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private funding flows. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO regulatory environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and its main funding sources. We conclude by discussing the implications of our results on the scholarship and practice of nonprofits using different framing appeals to increase fundraising at a time when traditional funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "8,031"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

Over the last three decades, governments across the world have sought to limit the work of nonprofits and nongovernmental organizations (NGOs)—a phenomenon known as "closing civic space" [@Carothers:2015; @DPR:2015]. Repressive governments have enacted laws creating barriers to advocacy, funding, and entry, adversely affecting the work of thousands of NGOs, which typically rely on funding from donors across the world. This is particularly the case in countries with weak institutions and poor governance, where international donors typically seek out NGOs to implement their projects, as direct aid transfers to recipient governments pose the risk of misuse and bureaucratic inefficiency [@Dietrich2013]. As a result of this global crackdown on civil society, however, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know there has been a negative impact on funding from official governmental donors, we know relatively little about how this crackdown affects private donors, in particular how it influences individual donors' willingness to engage in philanthropy.

Charitable giving to nonprofits working in international affairs has steadily increased over the past two decades, with a noticeable increase over the last decade. Between 2014 and 2016, giving to international affairs-focused issues increased by 14.1% [@givingusa2017report]. This amount is not insignificant—by 2016, total private giving to civil society organizations working in international affairs had increased to \$22.03 billion [@givingusa:2017]. Most notably, a large portion of this growth is driven by individuals making small-scale donations. In 2017, individuals gave $286.7 billion, or approximately 70% of total giving to international affairs. This further increased to \$427.71 billion in 2018, 85.7% of which was through individuals [@givingusa:2019]. Much of the existing research on donations to international causes is restricted to elite, high net-worth donors [@USTrust2014], but these large amounts demonstrate that it is imperative to understand the motivations and attributes of more ordinary individual donors. What factors affect individual donor preferences on giving to international causes? More specifically, how does information about repressive regulatory environments abroad affect donors' willingness to donate internationally, and how does knowledge of legal crackdown interact with other organizational attributes?

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@CrosonShang:2011; @Tremblay-BoirePrakash:2017]. We use an survey experiment to assess how different structural and organizational-level heuristics affect individual donors' willingness to donate. We find that learning about more repressive NGO regulatory environments increases generosity in that already likely donors are willing to donate substantially more to legally restricted NGOs.

This generosity persists when mediated by two important organization-level heuristics: issue area and funding source. Learning about crackdown increases individuals' willingness to donate to privately funded nonprofits. This may be because the survival of the NGO appears to be in question and the donor's assessment might be that their donation actually makes a difference. These results are particularly substantive for privately-funded human rights NGOs facing crackdown, with donors showing an increased willingness to not just donate to them, but also to donate more to them. While not unsurprising, this does suggest that private donors to human rights NGOs likely know that their work is always challenging to host governments, and that such organizations need greater support when facing stricter regulatory environments.

Our study makes a significant contribution to research on individual giving and indicates the promise and limits of different framing appeals on individuals' willingness to donate. First, while factors shaping individual giving to domestic causes have been thoroughly explored in existing literature, the dynamics of international giving and individual donor preferences are less explored.[^exceptions] Understanding the factors driving individual giving is particularly important when when looking at international issues, as philanthropy to international causes can be harder to motivate because of the number of recipients is larger and further removed from the donor [@CasaleBaumann:2015, 100].

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. But in the era of closing civic space, these groups may need to tailor their fundraising strategies to individual donors. Framing, or the process through which actors present information to influence perceptions of behavior, is particularly important in this regard [@ChongDruckman:2007]. However, we lack systematic studies of which frames may be effective when making appeals to donors considering international philanthropy. Our study examines the impact of political frames such as regulatory crackdown, and how these frames interact with organizational attributes to signal NGO deservingness. Accordingly, our results can help international nonprofits frame their appeals for funding, especially when facing restricted legal space abroad.

@@

Thousands of international NGOs in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though we focus on private funding in this article, it is essential to differentiate between the two sources. Public aid, or more traditional donor aid, channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds come from foundations, private corporations, and individuals. We are specifically interested in individual-level determinants to private international philanthropy. Private giving to nonprofits working in international affairs has increased consistently over the past decade.[^internationalaffairs] Moreover, donations from individuals have far outstripped foundation and corporate giving—in 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion by corporations. While foundations are often seen as the primary funders of international charities, donations from foundations accounted for only 18% of all charitable giving in 2018—and grants from the Bill and Melinda Gates Foundation accounted for more than half of all foundation giving [@givingusa:2019]. Individual donors play an incredibly important—and underexplored—role in funding international NGOs.

Given these substantive amounts, understanding individual-level determinants of giving to international nonprofits is important to investigate as individual private donors may not necessarily mirror the preferences of official aid agencies. While there is evidence that citizens generally support the goals of aid agencies [@MilnerTingley2010], surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, both foreign aid and democracy aid are often considered unresponsive to political developments in recipient countries [@Carothers:2015] and public aid decisions—especially regarding final recipients and amounts—are not easily accessible. There is also heterogeneity in how donor governments allocate funds. For instance, unlike established donors like the United States and other OECD countries, emerging donor governments tend to demand fewer human rights, governance, or environmental preconditions [@fenglerkharas].

More recent research on the allocation of private aid flows has also found little support for the idea that individual donor behavior mimics that of official donors. @desaikharas find that unlike official donors who tend to channel funds to countries with better institutional quality or as a reward for governmental performance [@Bermeo:2011; @DietrichWright:2015], individual donors do not use the same performance-based metrics. Rather, individuals are more inclined to donate to groups in countries experiencing a humanitarian crisis—in particular, crises following natural disasters. @desaikharas conclude that compared to donor agencies, "private donors respond to different project and country characteristics" (p. 517) Given these above reasons, the average individual donor will not follow the same decision-making logic as donor governments, and individual international philanthropy should therefore be investigated on its own terms.

Understanding the motivations of individual private donors is also essential because private funding flows towards nonprofits working internationally appear to not be as adversely impacted as public funding flows in the face of the global crackdown on NGOs. While anti-NGO restrictions have reduced official aid to repressive countries [@DupuyPrakash:2018], private foundations have continued to channel funds to countries with unfriendly legal environments [@Foundation-Center:2018]. Accordingly, it is imperative to understand whether individual donors mirror the trajectory of foundations when giving internationally, especially to organizations facing legal crackdown abroad.

# Key drivers of individual-level philanthropy

@@

Prior research has found that those who participate in a variety of associations and build social capital are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not), were both more likely to make charitable donations and make bigger donations, since volunteering exposed individuals to the need for charitable donations and made philanthropy more compelling [@hossainlamb2012]. An individual's social networks also play an important in deciding to donate. Looking at relational configurations in social networks, @HerzogYang:2018 find that having both a giver and a solicitor in one's social network increase the likelihood of that individual's charitable giving.

Individual experiences such as access to higher education and increased religiosity can also influence philanthropic giving. Higher education fosters prosocial motivations and brings people into particular social networks that entail a higher level of solicitations [@BekkersWiepking:2011]. Education also increases abstract thinking [@wiepkingmaas, 1978], which is particularly important for donating to more distant international relief organizations than donating to local, more visible nonprofits [@Micklewright:2009]. Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attend religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform to group and community standards [@wiepking2010]. Overall, individuals with higher levels of income, education, and greater religious proclivities have been shown to not only be more likely to give internationally, but also to give higher amounts [@Rajan:2009; @CasaleBaumann:2015].

Finally, information about organizational characteristics and third-party certifications can also impact levels of individual philanthropy, through results are mixed. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. The effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010] and that ratings by charity watchdogs do not really affect donor support for these nonprofits [@SzperPrakash:2011].

While this robust literature looks at the drivers of individual-level philanthropy, it overwhelmingly focusses on giving to local rather than international causes. This domestic focus is warranted, though: on average, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad" [@Tremblay-BoirePrakash:2017, 644]. Accordingly, there is a lack of data and theory regarding international philanthropy. Most research on individual donors to international causes has been restricted to elite, high net-worth donors (those with more than $200,000 in annual income or $1 million in assets) [@USTrust2014]. Given the increasing importance of more ordinary individual donations for international NGOs, it is crucial to better understand what motivates private donors to engage in internationally-focused philanthropy.

Individuals rely on heuristics that differ from the benchmarks used by large governmental agencies and private foundations. These signals, we argue, can be even more important in giving to international causes. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, donors respond to a host of heuristics when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. Nonprofits fundraising internationally may therefore rely framing to control perceptions. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

@@

We argue that donors perceive legal crackdowns differently from organization-level heuristics—such as NGO missions, visions, values—because donors may be able to influence these latter attributes, at least to some extent. Conversely, donors can do little to change the domestic political environment of another country. Legal crackdowns abroad can signal to donors that the government may eventually target other NGOs as well, which can increase donor urgency to support legally besieged NGOs, even if donors are not completely satisfied with some organization-level features of a group. As such, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization and increase the amount donated.

@@

The *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences, can also serve as an organization-level heuristic.[^targeting] NGOs that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. Previous research has shown that that public donors do indeed differentiate between NGO issue areas, especially when these organizations are working in repressive contexts. In particular, donor agencies in the Organisation for Economic Cooperation and Development (OECD) respond to repressive NGO environments, especially those containing barriers to advocacy, by decreasing funds for groups working on politically sensitive causes such as anti-corruption initiatives, elections, human rights, legal reform and security sector reform, and instead increasing funding for groups working on relatively tamer causes such as health, education and agriculture [@ChaudhryHeiss:2018].

Issue area contentiousness directly influences legal crackdowns, however, and confounds the heuristic effect of crackdowns on donor behavior (see @fig:causal-path). NGOs working on more contentious human rights issues can be challenging to regimes [@Heiss:2017]. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. There is evidence that this a concern for private donors. In a survey of women’s NGOs across the globe conducted by Mama Cash and Urgent Action Fund, many NGOs argued that private donors (especially foundations) were withdrawing from funding sensitive issues in repressive contexts, which in turn hurt women and trans rights organizations. Some private funders, these groups argued, were also more likely to fund well-established groups that were more likely to survive crackdown.

However, compared to our existing knowledge on public donors and foundation donors, we lack knowledge about how individual donors respond to frames emphasizing one nonprofit issue area over another, especially a contentious area over a tamer one. NGOs working on tamer issues such as health, education, disaster relief, and humanitarian aid can often frame their activities as in need of greater support—the perception of need, whether through personal experiences or media framing. This may strongly influence individual decisions to donate to relief efforts [@wiepking2010]. Individual donors may believe that it is more difficult to improve human rights in the host country and subsequently shy away from donating to NGOs working on contentious causes.

@@

The final heuristic we examine is the *source of NGO funding*. Government grants to nonprofits can often crowd out private funding, though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991; @Heutel:2014]. Donors may feel less inclined to contribute to NGOs receiving funds from agencies like USAID because such funding can signal a lack of independence from the government. Organizations that receive substantial funding from their home governments frequently avoid programming that would question the donor government even if such programming is in line with their mission [@Stroup:2012]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help and that their marginal donation would make a noticeable difference.

@@

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Donors do not always know about anti-NGO laws in host countries or who funds an NGO when they make their contributions. Moreover, the average donor will not know exactly what "legal crackdown" entails for any given organization (i.e. does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguous definition is not a central concern as the heuristics we explore here are explicitly simplified shortcuts to donor information-gathering. Knowing that an organization works with human rights issues (even if donors are not familiar with specific kinds of rights) should be enough to trigger the issue-based deservingess heuristic. Similarly, knowing that an NGO faces legal trouble should be enough to trigger the structural deservingness heuristic. Moreover, while many NGOs and nonprofits make information about their funding sources easily accessible through annual reports, donors often do not take the time to track down that information. In each of these hypotheses, we assume that donors will not spend additional effort to research the exact details concerning crackdown, issue area, and funding sources. Instead, donors rely on signals in basic information about the organization to shape their preferences and behavior.

@@

To measure the effect of structural and organization-level heuristics on individual donor preferences, we use an experiment to vary the different frames donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs tailor their fundraising strategies for private individuals. Prior to launching the experiment, we preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at [https://osf.io/dx973/](https://osf.io/dx973/). The online appendix also includes the full text of the survey experiment, as well as details about our sample demographics, balance across experimental conditions, CONSORT diagram, and Bayesian priors. The analysis is fully reproducible using code and data available at ANONYMIZED.

@@

We presented participants with a short paragraph with three manipulated frames, each highlighting different factors that might influence patterns of private philanthropy: (1) crackdown/no crackdown, (2) humanitarian assistance/human rights, and (3) government/private funding. We used the International Rescue Committee (IRC) as our example NGO because it fits within each possible frame. IRC provides humanitarian assistance and engages in human rights advocacy for refugees, more than a quarter of IRC's income comes from both government grants and private donations, and it works in countries with and without anti-NGO laws. We use a 2×2×2 between-subject factorial design with participants randomly assigned to one of eight versions of the following vignette:

@@

[^internationalaffairs]: According to Giving USA, international affairs includes nonprofits working in international development, international relief services, disaster relief, international human rights, international peace and security, foreign policy research and analysis, and international exchange programs.

@@

[^personalframes]: Personal frames humanize victims and entice action by invoking an emotional reaction. For instance. people are more be generous when donating to a human rights campaign with a single victim, rather than a statistical group of victims [@Slovic:2007].

>>>>

---
title: "Dynamics of international giving: Exploring individual donor preferences"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Lewis \& Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  Government restrictions on non-governmental organizations (NGOs) have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private philanthropy. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and its main funding sources. We discuss the implications of our results on how nonprofits can use different framing appeals to increase fundraising at a time when traditional public donor funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "8,031"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression, international giving, donor choice
---

@@

Over the last three decades, governments across the world have sought to limit the work of nonprofits and nongovernmental organizations (NGOs)—a phenomenon known as "closing civic space" [@Carothers:2015; @DPR:2016]. Repressive governments have enacted laws creating barriers to advocacy and funding for a variety of NGOs. The adverse effects of these laws are particularly felt in countries with weak institutions and poor governance. International donors directing aid to these countries typically seek out NGOs to implement their projects, as direct aid transfers to recipient governments pose the risk of misuse and bureaucratic inefficiency [@Dietrich2013]. However, official aid flows to NGOs have decreased considerably in countries with repressive NGO environments [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know there has been a negative impact on funding from official governmental donors, we know relatively little about how this crackdown affects private donors, in particular how it influences individual donors' willingness to donate internationally.

Charitable giving to nonprofits working in international affairs has steadily increased over the past two decades. Between 2014 and 2016, giving to international affairs-focused issues increased by 14.1% [@givingusa2017report]. This amount is not insignificant—by 2016, total private giving to civil society organizations working in international affairs had increased to \$22.03 billion [@givingusa:2017]. Most notably, a large portion of this growth is driven by individuals making small-scale donations. In 2017, individuals gave $286.7 billion, or approximately 70% of total giving to international affairs. This further increased to \$427.71 billion in 2018, 85.7% of which was through individuals [@givingusa:2019]. While a lot is known about individuals' motivations for giving locally [@BekkersWiepking:2011], much less is known about motivations to give internationally and these large amounts show the urgency of understanding this phenomenon. What factors affect individual donor preferences on giving to international causes? More specifically, how does information about repressive NGO environments abroad affect donors' willingness to donate internationally, and how does knowledge of legal crackdown interact with other organizational attributes?

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@CrosonShang:2011; @Tremblay-BoirePrakash:2017]. We use a survey experiment fielded in the U.S. to assess how different structural and organizational-level heuristics affect these donors' willingness to donate. We find that learning about more repressive NGO regulatory environments increases generosity in that already likely donors are willing to donate substantially more to legally restricted NGOs.

This generosity persists when mediated by two important organization-level heuristics: issue area and funding source. Learning about crackdown increases individuals' willingness to donate to privately funded nonprofits. This may be because the survival of the NGO appears to be in question and the donor's assessment might be that their donation actually makes a difference. These results are particularly substantive for privately-funded human rights NGOs facing crackdown, with donors showing an increased willingness to not just donate to them, but also to donate *more* to them. While not unsurprising, this does suggest that private donors to human rights NGOs likely know that their work is always challenging to host governments, and that such organizations need greater support when facing hostile environments.

Our study makes a significant contribution to research on individual giving and indicates the promise and limits of different framing appeals on individuals' willingness to donate internationally. While factors shaping individual giving to domestic causes have been thoroughly explored in existing literature, the dynamics of international giving and individual donor preferences are less explored.[^exceptions] Understanding the latter is important as philanthropy to international causes can be harder to motivate because the number of recipients is larger and further removed from the donor [@CasaleBaumann:2015, 100].

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. But in the era of closing civic space, these groups may need to tailor their fundraising strategies to individual donors. Framing, or the process through which actors present information to influence perceptions of behavior, is particularly important in this regard [@ChongDruckman:2007]. However, we lack systematic studies of which frames may be effective when making appeals to donors considering international philanthropy. Our study examines the impact of political frames such as regulatory crackdown, and how this frame interacts with organizational attributes to signal NGO deservingness. Accordingly, our results can help international nonprofits frame their appeals for funding, especially when facing restricted legal space abroad.

@@

Thousands of international NGOs in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though we focus on private funding in this article, it is essential to differentiate between the two sources. Public aid, or more traditional donor aid, channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds come from foundations, corporations, and individuals. We are specifically interested in individual-level determinants of private international philanthropy.

Private giving to nonprofits working in international affairs has increased consistently over the past decade.[^internationalaffairs] Moreover, donations from individuals have far outstripped foundation and corporate giving—in 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion by corporations. While foundations are often seen as the primary funders of international nonprofits, donations from foundations accounted for only 18% of all charitable giving in 2018—and grants from the Bill and Melinda Gates Foundation accounted for more than half of all foundation giving [@givingusa:2019]. Individual donors thus play an incredibly important—and underexplored—role in funding international NGOs.

Given these substantive amounts, understanding individual motivations to donate internationally is important to investigate as individual donor preferences may not necessarily mirror the preferences of official aid agencies. While there is evidence that citizens generally support the goals of aid agencies [@MilnerTingley2010], surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, both foreign aid and democracy aid are often considered unresponsive to political developments in recipient countries [@Carothers:2015] and public aid decisions—especially regarding final recipients and amounts—are not easily accessible. There is also heterogeneity in how donor governments allocate funds. For instance, unlike established donors such as the United States and other Organisation for Economic Cooperation and Development (OECD) countries, emerging donor governments tend to demand fewer human rights, governance, or environmental preconditions [@fenglerkharas].

Recent experimental evidence also finds little support for the idea that individual donor behavior mimics that of official donors. @desaikharas find that unlike official donors who tend to channel funds to countries with better institutional quality or as a reward for governmental performance [@Bermeo:2011; @DietrichWright:2015], individual donors do not use the same performance-based metrics. Rather, individuals are more inclined to donate to NGOs in countries experiencing a humanitarian crisis—in particular, crises following natural disasters. @desaikharas conclude that compared to donor agencies, "private donors respond to different project and country characteristics." (p. 517) International philanthropy by individuals should therefore be investigated on its own terms.

Looking at different funding flows, it is also notable that private funding towards nonprofits working internationally appear to not be as adversely impacted as public funding flows in the face of the global crackdown on NGOs. While anti-NGO restrictions have reduced official aid to repressive countries [@DupuyPrakash:2018], private foundations have continued to channel funds to countries with unfriendly legal environments [@Foundation-Center:2018]. Accordingly, it is imperative to understand whether individual donors mirror the trajectory of foundations when giving internationally, especially to organizations facing legal crackdown abroad.

## Key drivers of individual-level philanthropy

@@

Prior research has found that those who participate in a variety of associations and build social capital are more likely to make charitable donations as well as make larger donations [@hossainlamb2017, @hossainlamb2012]. An individual's social networks also play an important in deciding to donate. Looking at relational configurations in social networks, @HerzogYang:2018 find that having both a giver and a solicitor in one's social network increased the likelihood of that individual's charitable giving.

Individual experiences such as access to higher education and increased religiosity can also influence giving [@wiepking2010]. Higher education fosters pro-social motivations and brings people into social networks that entail a higher level of solicitations [@BekkersWiepking:2011]. Education also increases abstract thinking [@wiepkingmaas, 1978], which is important for donating to distant international relief organizations than donating to local, more visible nonprofits [@Micklewright:2009]. Overall, individuals with higher levels of income, education, and greater religious proclivities have been shown to not only be more likely to give internationally, but also to give higher amounts [@Rajan:2009; @CasaleBaumann:2015].

Finally, information about organizational characteristics and third-party certifications can also impact levels of individual philanthropy, though results are mixed. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. The effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010] and that ratings by charity watchdogs do not really affect donor support for these nonprofits [@SzperPrakash:2011].

Even though these above factors affect individuals' motivations to donate internationally, extant literature on international giving focuses on donating during crises and emergencies. However, as @Micklewright:2009 point out, motivations for donating to ongoing international causes differ from donating money to natural disasters or emergency international causes. Under ordinary circumstances, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad" [@Tremblay-BoirePrakash:2017, 644]. However, we lack adequate data and theory regarding international philanthropy. Most research on individual donors to international causes has been restricted to elite, high net-worth donors -- those with more than $200,000 in annual income or $1 million in assets [@USTrust2014]. However, given the increasing funds channeled by individuals to international nonprofits, it is crucial to understand what motivates these donors to donate.

Individuals rely on heuristics that differ from the benchmarks used by large governmental agencies and private foundations. These signals, we argue, can be even more important in giving to international causes. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, donors respond to a host of heuristics when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. Nonprofits fundraising internationally may therefore rely on framing to control perceptions. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

@@

We argue that donors perceive legal crackdowns differently from organization-level heuristics—such as NGO missions, visions, values—because donors may be able to influence these latter attributes, at least to some extent. Conversely, donors can do little to change the domestic political environment of another country. Legal crackdowns abroad can signal to donors that the government may eventually target other NGOs as well, which can increase donor urgency to support legally besieged groups, even if donors are not completely satisfied with some organization-level features of a group. As such, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization and increase the amount donated.

@@

The *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences, can also serve as an organization-level heuristic.[^targeting] NGOs that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. Previous research has shown that public donors do indeed differentiate between NGO issue areas, especially when these organizations are working in repressive contexts. In particular, donor agencies in the Organisation for Economic Cooperation and Development (OECD) respond to repressive NGO environments, especially those containing barriers to advocacy, by decreasing funds for groups working on politically sensitive causes such as anti-corruption initiatives, elections, human rights, legal reform and security sector reform, and instead increase funding for groups working on relatively tamer causes such as health, education and agriculture [@ChaudhryHeiss:2018].

Issue area contentiousness directly influences legal crackdowns, however, and confounds the heuristic effect of crackdowns on donor behavior (see @fig:causal-path). NGOs working on more contentious human rights issues can be challenging to regimes [@Heiss:2017]. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. There is evidence that this a concern for private donors. In a survey of women’s NGOs across the globe conducted by Mama Cash and Urgent Action Fund, many groups argued that private donors (especially foundations) were withdrawing from funding sensitive issues in repressive contexts, which in turn hurt women and trans rights organizations. Some private funders, these groups argued, were also more likely to fund well-established groups that were more likely to survive crackdown [@mamacash].

However, compared to our existing knowledge on public donors and foundation donors, we lack knowledge about how individuals respond to frames emphasizing one nonprofit issue area over another, especially a contentious area over a tamer one. NGOs working on tamer issues such as health, education, disaster relief, and humanitarian aid can often frame their activities as in need of greater support. This may strongly influence individual decisions to donate to relief efforts [@wiepking2010]. On the flip side, ndividual donors may believe that it is more difficult to improve human rights in the host country and subsequently shy away from donating to NGOs working on contentious causes.

@@

The final heuristic we examine is the *source of NGO funding*. Government grants to nonprofits can often crowd out private funding, though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991; @Heutel:2014]. Donors may feel less inclined to contribute to NGOs receiving funds from agencies like United States Agency for International Development (USAID) because such funding can signal a lack of independence from the government. Organizations that receive substantial funding from their home governments frequently avoid programming that would question the donor government even if such programming is in line with their mission [@Stroup:2012]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help and that their marginal donation would make a noticeable difference.

@@

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Moreover, the average donor will not know exactly what "legal crackdown" entails for any given organization (i.e. does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguity is not a central concern as the heuristics we explore here are explicitly simplified shortcuts to donor information-gathering. Knowing that an organization works with human rights issues (even if donors are not familiar with specific kinds of rights) should be enough to trigger the issue-based deservingness heuristic. Similarly, knowing that an NGO faces legal trouble should be enough to trigger the structural deservingness heuristic. Moreover, while many NGOs make information about their operations accessible through annual reports, it is time-consuming to track down that information. In each of these hypotheses, we assume that donors will not spend additional effort to research the exact details concerning crackdown, issue area, and funding sources. Instead, donors rely on signals and heuristics about the organization to shape their preferences and behavior [@SzperPrakash:2011].

@@

To measure the effect of structural and organization-level heuristics on individual donor preferences, we use an experiment to vary the different frames donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs tailor their fundraising strategies for individuals. Prior to launching the experiment, we preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at [https://osf.io/dx973/](https://osf.io/dx973/). The online appendix also includes the full text of the survey experiment, as well as details about our sample demographics, balance across experimental conditions, CONSORT diagram, and Bayesian priors. The analysis is fully reproducible using code and data available at ANONYMIZED.

@@

We presented participants with a short paragraph with three manipulated frames, each highlighting different factors that might influence patterns of private philanthropy: (1) crackdown/no crackdown, (2) humanitarian assistance/human rights, and (3) government/private funding. We used the International Rescue Committee (IRC) as our example NGO because it fits within each possible frame and as such, involves no deception towards respondents. IRC provides humanitarian assistance and engages in human rights advocacy for refugees, more than a quarter of IRC's income comes from both government grants and private donations, and it works in countries with and without anti-NGO laws. We use a 2×2×2 between-subject factorial design with participants randomly assigned to one of eight versions of the following vignette:

@@

[^internationalaffairs]: According to Giving USA, the category of `international affairs' includes nonprofits working in international development, international relief services, disaster relief, international human rights, international peace and security, foreign policy research and analysis, and international exchange programs.

@@

[^personalframes]: Personal frames humanize victims and entice action by invoking an emotional reaction. For instance, people are more be generous when donating to a human rights campaign with a single victim, rather than a statistical group of victims [@Slovic:2007].
